---
title: "Assignment 3-3"
author: "Allyson Busch"
date: "September 14, 2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

For this week's assignment we're going to use Dodgers Major League Baseball data from 2012. I would like you to determine what night would be best to run a promotion to increase attendance. You will want to use TRAIN. As a reminder, the training set is the data we fit our model on. Use a combination of R and Python to accomplish this assignment. You should be able to demonstrate the following: box plots, scatter plots, regression model.

```{r loading}

library('magrittr')

options(plot.height= 3, plot.width = 3)

dodgers <- read.csv('dodgers.csv')

```

## Exploratory Analysis


```{r checking data}

head(dodgers)

```

```{r continue}

dodgers$day_of_week <- factor(dodgers$day_of_week, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

dodgers$month <- factor(dodgers$month, levels = c("APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT"))

summary(dodgers)

```

THere are promotions that range from April to August, that range in type from Cap, Shirt, Fireworks, and Bobbleheads. The data is divided to the seven days of the week. The maximum attendance is 56,000. 

## Boxplot

```{r boxplot}

# attendance by day

ggplot2::ggplot(dodgers, ggplot2::aes(x=day_of_week, y = attend/1000)) + 
  ggplot2::geom_boxplot(color='blue', fill='darkblue', alpha = 0.2) +
  ggplot2:: labs(title="Attendance by Day of the Week", x = 'Day of Week', y = 'Attendance (in thousands)') + 
  ggplot2::theme_bw() + 
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))

```

Based on the boxplots, Tuesdays have the highest attendance since the median is the highest and mondays are the lowest attendance day. 

```{r attendance by month}

ggplot2::ggplot(dodgers, ggplot2::aes(x=month, y=attend/1000)) + 
  ggplot2::geom_boxplot(color="blue", fill="dark blue", alpha = 0.5) + 
  ggplot2::labs(title = 'Attendance by Month', x = 'Month', y = 'Attendance (in thousands)') + 
  ggplot2::theme_bw() + 
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))

```

Based on the above box plot, June is the highest attendance month because the median is the highest. October is the lowest month.

``` {r weather}

ggplot2::ggplot(dodgers, ggplot2::aes(x = temp, y = attend/1000, colors = fireworks)) + 
  ggplot2::geom_point() + 
  ggplot2::facet_wrap(day_night~skies) + 
  ggplot2::labs(title = 'Attendance by Temperature, Time and Skies', x = 'Temperature', y = 'Attendance (Thousands)')

```

Based on the above scatterplots, there was only one day game played under a cloudy sky, so we don't have enough information to make good observations on that. Most games were played at night with a clear sky, and it had higher attendance. 

```{r opponents}

ggplot2::ggplot(dodgers, ggplot2::aes(x = attend/1000, y = opponent, color = day_night)) + 
  ggplot2::geom_point() + 
  ggplot2::labs(title = 'Attendance by Opponent', x = 'Attendance (in thousands)', y = 'Opponent') + 
  ggplot2::theme(plot.title = ggplot2::element_text(lineheight = 3, face= 'bold', color = 'black', size = 10))

```

## Predictive Model

```{r model}

dodger.model <- {attend ~ month + day_of_week + fireworks}

set.seed(123)
traintest_key <- c(rep(1, trunc((2/3)*nrow(dodgers))), rep(2, trunc((1/3)*nrow(dodgers))))

dodgers$train_test <- sample(traintest_key)

dodgers$train_test <- factor(dodgers$train_test, levels = c(1,2), labels = c("TRAIN", "TEST"))

train <- subset(dodgers, train_test == "TRAIN")
test <- subset(dodgers, train_test == "TEST")

model.fit.train <- lm(dodger.model, data = train)

train$attend_pred <- predict(model.fit.train)

test$attend_pred <- predict(model.fit.train, newdata = test)

accuracy <- round(cor(test$attend, test$attend_pred)^2, digits = 3)

print(accuracy)

```

``` {r estimates}

#estimating the attendance based on the full data set

dodgers_predicted <- train %>% dplyr::bind_rows(test)
model.fit.all <- lm(dodger.model, data = dodgers)

print(summary(model.fit.all))

```

Since the coefficient t-value is a measure of how many standard deviations the coefficient estimate is, and ours is 12.058, fireworks are not a good predictor. With the R-squared being 0.449, only 44.9% of the variance found in the variable can be explained by the predictor variable. 

Because of this low number, I decided to run it again with the Bobblehead promotion. 

```{r model part 2}

dodgers2 <- dodgers

dodger.model2 <- {attend ~ month + day_of_week + bobblehead}

set.seed(1234)
traintest_key2 <- c(rep(1, trunc((2/3)*nrow(dodgers2))), rep(2, trunc((1/3)*nrow(dodgers2))))

dodgers2$train_test <- sample(traintest_key)

dodgers2$train_test <- factor(dodgers2$train_test, levels = c(1,2), labels = c("TRAIN", "TEST"))

train2 <- subset(dodgers2, train_test == "TRAIN")
test2 <- subset(dodgers2, train_test == "TEST")

model.fit.train2 <- lm(dodger.model2, data = train2)

train2$attend_pred <- predict(model.fit.train2)

test2$attend_pred <- predict(model.fit.train2, newdata = test2)

accuracy2 <- round(cor(test2$attend, test2$attend_pred)^2, digits = 3)

print(accuracy2)

```

This has a much higher accuracy than the fireworks predictor. 

``` {r estimates}

dodgers_predicted2 <- train2 %>% dplyr::bind_rows(test2)
model.fit.all2 <- lm(dodger.model2, data = dodgers2)

print(summary(model.fit.all2))


```

The coefficient t value is still high for the intercept, but is low for most of the variables. The lowest was Thursday, being there was little deviation from guessing attendance on Thursdays. The p value is significantly low again, and the r-square accounts for 54.4% of the variance found in the response variable. 

The best night to run a promotion would likely be Thursday nights based on the regression above. 

